---
title: "RotaTiters analysis"
author: "Dr. Marina Antillon and Prof. Meagan Fitzpatrick"
date: "31 March 2026"
output:
  word_document: 
    toc: yes
  pdf_document: default
---

<!-- # How to use this Rmarkdown -->

<!-- This Rmarkdown document is meant to be used with Rstudio and it outputs a Word document. When you click the **Knit** button at the top left, a Word document will be generated that includes both content as well as the output of any embedded R code chunks within the document. -->

<!-- One may run individual chunks by hitting the green right-facing triangle on the top right of each chunk. -->

<!-- By default, the R code is excluded from the Rmarkdown, as given by the first line of code of the following chunk. Note that the `echo = FALSE` or `echo = TRUE` parameter can be added to each chunk specifically to prevent or dis/enable printing of the R code that generated the table or plot. -->

```{r setup, include=FALSE}

getwd()
# Whether to print R code
knitr::opts_chunk$set(echo = FALSE)
set.seed(123)

# Packages ----------------
# library("lubridate") # to work with dates
library("magrittr") # for the pipes: %>% 
library("ggplot2") # to make graphs
library("grid") # for annotation layer
library("readr") # To read csv's
library("tidyr") # tidy package for data; pivot definitely needs this
library("dplyr") # other tidy world things: mutate, summarize, group_by
library("stringr") # to deal with strings
library("cowplot") # multi-panel plots
library("ggpubr") # still not a stable version; nec for Jruskall wallis tests on graphs
library("MASS") # for the mvtnorm distribution
library("lmtest") # to look at linear model tests and (maybe diagnostics)
library("moments") # for the Jarque-Bera test for normality
library("mgcv") # for GAMs
library("caret") # for the leave-k-out validation that lets me choose splines' k parameter
library("scales") # to make nice-looking y-axes
library("flextable") # to output tables that work well in Word/PPT
library("modelsummary") # to have a neat summary of multiple models at once

# Helpful functions
fls = list.files("./functions/", pattern="^[fn]")
for (i in 1:length(fls)){source(paste0("./functions/", fls[i]))}

mean_ci = function(x){c(mean(x), quantile(x, c(0.025, 0.975)))}
med_iqr = function(x){quantile(x, c(0.5, 0.25, 0.75))}

# Create the "out" dir if not available
create_dir_if_necessary("./out/")

```

\newpage
# Objective of the analysis

This is an analysis of Rotavirus antibody titers in a cohort of Malian mothers who enrolled in a study of maternal flu vaccine in 2012 (https://clinicaltrials.gov/study/NCT01430689; https://doi.org/10.1016/S1473-3099(16)30054-8). In 2018-2021, 344 samples out of the original 4000+ samples were assayed for anti-Rotavirus IgA and IgG. Alongside, we will examine the cross-correlation between titers and incidence estimated for the same period (https://doi.org/10.1016/S2214-109X(19)30076-2; https://doi.org/10.1016/S2214-109X(19)30541-8). 

<!-- # Bring in the data -->

```{r calldata, include=F}
# Nothing in this chunk is shown in the word doc. That's the way it should be.

data = read_csv("./data/synthetic_rotavirus_titers-final.csv") %>%
    mutate(DOB=paste0("2012-",sprintf("%02d", month_num),"-",sprintf("%02d", day_num))) %>%
    mutate(DOB=base::as.Date(DOB)) %>% 
  rename(IgATiter=iga, IgGTiter=igg)

# Make a month variable
data = data %>% mutate(month = as.numeric(format(DOB, "%m")))
# it's easier if the DOB variable is just the number of days since Jan 1
data = data %>% mutate(dayofyear = as.numeric(base::as.Date(DOB) - base::as.Date("2012-01-01")))

# Uncomment this to save the data without the omitted individuals
# write.csv(data[data$IgATiter>8, c("iid", "mid")], "./data/motherIDsincluded.csv", row.names = F)
# write.csv(data[data$IgATiter<8, c("iid", "mid")], "./data/motherIDsexcluded.csv", row.names = F)

```

<!-- # Omitting readings of IgA<7.5 U/mL -->
<!-- The results of this section went into the methods supplement. -->

```{r omitlowIgA, include=F}

# Correlation IgA IgG ----------------

# plot(data$IgATiter, data$IgGTiter)

p1 = data %>% 
  ggplot(aes(x=IgATiter, y=IgGTiter)) + 
  geom_point() + 
  scale_x_log10(limits=c(5, NA)) + scale_y_log10(limits=c(5, NA)) +
  stat_cor(method = "pearson", label.x = 1, label.y = 4.2) + 
  ylab("IgA") + xlab(" IgG") + ggtitle("Correlation of IgA and IgG") +
  theme_bw()

p2 = data %>% filter(IgATiter>7.5) %>%
  ggplot(aes(x=IgATiter, y=IgGTiter)) + 
  geom_point() + 
  scale_x_log10(limits=c(5, NA)) + scale_y_log10(limits=c(5, NA)) +
  stat_cor(method = "pearson", label.x = 1, label.y = 4.2) + 
  ylab("IgA") + xlab(" IgG") + ggtitle("Correlation of IgA and IgG") +
  theme_bw()

pval_pearson_allIgA = with(data, cor.test(log2(IgATiter), log2(IgGTiter))) # parametric
pval_wilcox_allIgA = with(data, wilcox.test(IgATiter, IgGTiter)) # non-parametric

pval_pearson_subsetIgA = with((data %>% filter(IgATiter>7.5)), cor.test(log2(IgATiter), log2(IgGTiter))) # parametric
pval_wilcox_subsetIgA = with((data %>% filter(IgATiter>7.5)), wilcox.test(IgATiter, IgGTiter)) # non-parametric

# data %>% group_by(month) %>% 
#   dplyr::summarize(cor = cor(log2(IgATiter), log2(IgGTiter)),
#                    cor_lci = cor.test(log2(IgATiter), log2(IgGTiter))$conf.int[1],
#                    cor_hci = cor.test(log2(IgATiter), log2(IgGTiter))$conf.int[2],
#                    cor_pvalue = cor.test(log2(IgATiter), log2(IgGTiter))$p.value)

# Undetectable titers (<7.5) ------

IgAtoolow = data %>%
  group_by(month) %>%
  summarize(u7 = sum(IgATiter<7.51),
            o7 = sum(IgATiter>7.5)) %>% 
  mutate(month = month.name[month])

IgAtoolow_all = data %>%
  summarize(u7 = sum(IgATiter<7.51),
            o7 = sum(IgATiter>7.5)) %>%
  mutate(month = "Total") %>% relocate(month)

data_original = data
data = data %>% filter(IgATiter>7.5)

```

\newpage
# Description of the dataset

```{r desc-data, message=F, warning=F, fig.width=6, fig.height=3.5}

## GMTs ----------------------------------------------------

# Demographics
data %>%
  dplyr::summarise(Age = mean(mat_age),
                   Age_IQR = range_pretty(fivenum(mat_age)[c(2,4)]), 
                   Age_Range = range_pretty(range(mat_age)),
                   Birthweight = mean(bw),
                   Birthweight_IQR = range_pretty(fivenum(bw)[c(2,4)]), 
                   Birthweight_Range =range_pretty(range(bw)),
                   LBW = mean(bw<2500))

data %>% 
  dplyr::summarise(Boys = sum(baby_sex==1), Girls = sum(baby_sex==2))

data %>% dplyr::summarize(mean = mean(mat_age), 
                      IQR = range_pretty(quantile(mat_age, c(0.25, 0.75))),
                      Range = range_pretty(range(mat_age)))

t.test(data$mat_age, mu=26) # Mat Age
t.test(data$bw, mu=3217) # BW
binom.test(sum(data$bw<2500), length(data$month), p = 0.154, alternative = "two.sided") # BW, prop
binom.test(sum(data$baby_sex==1), length(data$month), p = 0.52, alternative = "two.sided") # Sex of the baby

data %>% 
  filter(mat_age<18) %>% 
  mutate(mat_age=floor(mat_age)) %>% 
  group_by(mat_age) %>% 
  dplyr::summarise(count = n())

data %>% 
  ggplot(aes(x=factor(month), y=bw)) + 
  geom_dotplot(binaxis="y",stackdir='center', binwidth = 50, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) + 
  scale_x_discrete(labels=month.abb) + 
  ylab("Birthweight") + xlab("") + # ggtitle("IgA by month of birth") + 
  theme_bw() + # theme(plot.title=element_text(hjust=0.5)) + 
  stat_compare_means(method = "kruskal.test", label.y = 4500, label.x = 2, 
                     label.sep=": ")

data %>% 
  ggplot(aes(x=factor(month), y=mat_age)) + 
  geom_dotplot(binaxis="y",stackdir='center', binwidth = 0.5, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) + 
  scale_x_discrete(labels=month.abb) + 
  ylab("Maternal age") + xlab("") + # ggtitle("IgA by month of birth") + 
  theme_bw() + # theme(plot.title=element_text(hjust=0.5)) + 
  stat_compare_means(method = "kruskal.test", label.y = 45, label.x = 2, 
                     label.sep=": ")

data %>% 
  ggplot(aes(x=factor(month), fill=factor(baby_sex))) + 
  geom_bar(position = "fill") + 
  scale_x_discrete(labels=month.abb) + 
  ylab("Percent boys") + xlab("") + # ggtitle("IgA by month of birth") + 
  theme_bw() + theme(legend.position="bottom")

contingency_table <- table(data$month, data$baby_sex)
chi_squared_test <- chisq.test(contingency_table)
chi_squared_test

```

# Possible confounders

```{r desc-conf, message=F, warning=F, fig.width=4, fig.height=3}

# Confounders vs the outcome
# cor.test(data$mat_age, log10(data$IgATiter))$estimate
# cor.test(data$mat_age, log10(data$IgATiter))$conf.int[1]
# cor.test(data$mat_age, log10(data$IgATiter))$conf.int[2]
outcomeA_matage = cor.test(data$mat_age, log10(data$IgATiter))$p.value

# cor.test(data$mat_age, log10(data$IgGTiter))$estimate
# cor.test(data$mat_age, log10(data$IgGTiter))$conf.int[1]
# cor.test(data$mat_age, log10(data$IgGTiter))$conf.int[2]
outcomeG_matage = cor.test(data$mat_age, log10(data$IgGTiter))$p.value

# cor.test(data$bw, log10(data$IgATiter))$estimate
# cor.test(data$bw, log10(data$IgATiter))$conf.int[1]
# cor.test(data$bw, log10(data$IgATiter))$conf.int[2]
outcomeA_bw = cor.test(data$bw, log10(data$IgATiter))$p.value

# cor.test(data$bw, log10(data$IgGTiter))$estimate
# cor.test(data$bw, log10(data$IgGTiter))$conf.int[1]
# cor.test(data$bw, log10(data$IgGTiter))$conf.int[2]
outcomeG_bw = cor.test(data$bw, log10(data$IgGTiter))$p.value

outcomeA_sex = kruskal.test(log10(data$IgATiter), factor(data$baby_sex))$p.value
outcomeG_sex = kruskal.test(log10(data$IgGTiter), factor(data$baby_sex))$p.value

# Exposures vs the outcome
exposure_matage = kruskal.test(data$mat_age, factor(data$month))$p.value
exposure_bw = kruskal.test(data$bw, factor(data$month))$p.value
exposure_sex = chisq.test(data$baby_sex, data$month)$p.value

confounders_tbl = data.frame(Confounder = c("Maternal Age", "Birthweight", "Infant Sex"),
                    IgA = c(outcomeA_matage, outcomeA_bw, outcomeA_sex),
                    IgG = c(outcomeG_matage, outcomeG_bw, outcomeG_sex),
                    Month = c(exposure_matage, exposure_bw, exposure_sex))

confounders_tbl %>%
  as_flextable(show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  align(j=2:4, align = "center", part = "header") %>% 
  align(j=2:4, align = "center", part = "body") %>% 
  add_header_row(values = c("", "Outcomes", "Exposure"), colwidths = c(1, 2, 1), top = TRUE) %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header") %>%
  colformat_double(j=2:4, digits = 3)
  
```

# Analysis: descriptive stats and CCF

```{r desc-mob, message=F, warning=F, fig.width=9, fig.height=7}

# By month
GMT_summaries1 = data %>% 
  group_by(month) %>% 
  dplyr::summarize(N = n(), 
                   IgA_med = ci_string_dec(quantile(IgATiter, c(0.5, 0.25, 0.75)), 0), 
                   IgA_GMT = ci_string_dec(c(GMT(IgATiter), GMT_CI_se(IgATiter)), 0), 
                   IgG_med = ci_string_dec(quantile(IgGTiter, c(0.5, 0.25, 0.75)), 0), 
                   IgG_GMT = ci_string_dec(c(GMT(IgGTiter), GMT_CI_se(IgGTiter)), 0))

GMT_summaries1_graph = data %>% 
  group_by(month) %>% 
  dplyr::summarize(N = n(), 
                   IgA_med = quantile(IgATiter, 0.5), 
                   IgA_q25 = quantile(IgATiter, 0.25), 
                   IgA_q75 = quantile(IgATiter, 0.75), 
                   IgA_GMT = GMT(IgATiter), 
                   IgA_GMTlci = GMT_CI_se_low(IgATiter), 
                   IgA_GMTuci = GMT_CI_se_high(IgATiter), 
                   IgG_med = quantile(IgGTiter, 0.5), 
                   IgG_q25 = quantile(IgGTiter, 0.25), 
                   IgG_q75 = quantile(IgGTiter, 0.75), 
                   IgG_GMT = GMT(IgGTiter), 
                   IgG_GMTlci = GMT_CI_se_low(IgGTiter), 
                   IgG_GMTuci = GMT_CI_se_high(IgGTiter))

# Might move this to the supplement or simply combine it with the boxplot.
GMT_summaries1_ggplot = GMT_summaries1_graph %>%
  pivot_longer(cols = starts_with(c("IgA", "IgG")),
               names_to = c("Ab", "measure"),
               names_pattern = "(.*)_(.*)") %>%
  pivot_wider(names_from=measure, values_from = value) %>%
  ggplot(aes(x=as.factor(month))) +
    geom_point(aes(y=med), color="red") +
    geom_errorbar(aes(ymin=q25, ymax=q75), color="darkred", linewidth=3) +
    geom_point(aes(y=GMT), color="blue") +
    geom_errorbar(aes(ymin=GMTlci, ymax=GMTuci), color="blue", linewidth=2) +
    facet_wrap(Ab~., scale="free_y", nrow=2) +
    scale_y_log10() +
    scale_x_discrete(label=month.abb) +
    xlab(NULL) +
    ylab("U/mL (log scale)") +
    theme_bw()

GMT_summaries_all = data %>% # (whole year)
  dplyr::summarize(N = n(), 
                   IgA_med = ci_string_dec(quantile(IgATiter, c(0.5, 0.25, 0.75)), 0), 
                   IgA_GMT = ci_string_dec(c(GMT(IgATiter), GMT_CI_se(IgATiter)), 0), 
                   IgG_med = ci_string_dec(quantile(IgGTiter, c(0.5, 0.25, 0.75)), 0), 
                   IgG_GMT = ci_string_dec(c(GMT(IgGTiter), GMT_CI_se(IgGTiter)), 0))

pval_pearson_allIgA = with(data, cor.test(log2(IgATiter), log2(IgGTiter))) # parametric
pval_wilcox_allIgA = with(data, wilcox.test(IgATiter, IgGTiter)) # non-parametric

GMT_summaries1 %>% 
  mutate(Month = month.name[month]) %>%
  relocate(Month) %>% 
  dplyr::select(-month) %>%
  bind_rows((GMT_summaries_all %>% mutate(Month = "All") %>% relocate(Month))) %>%
  as_flextable(max_row = 13, show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(2, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("Month", "N", "GMT (95% CI)", "Median (IQR)", "GMT (95% CI)", "Median (IQR)")) %>%
  align(align = "center", part = "header") %>% 
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header") %>%
  bg(i=13, bg = "gray80", part = "body") %>% 
  add_footer_lines(value=c("Correlation between IgA and IgG:",
                   paste0("Parametric test: coefficient: ", my_smallvalue(pval_pearson_allIgA$estimate, 3), " p", my_smallvalue(pval_pearson_allIgA$p.value, 3)),
                   paste0("Non-parametric test: coefficient: NA", " p", my_smallvalue(pval_wilcox_allIgA$p.value, 3))),
           top=FALSE)

## Draw Titer Patterns ------------------------------------
### IgA ----------------------------------------------------

BPd_IgA = ggplot(data, aes(x=DOB, y=IgATiter)) + 
  geom_point() + 
  geom_smooth(method=gam, formula=y ~ s(x, bs="cc", k=12)) + 
  scale_y_log10(limits=c(10, 12000)) + 
  scale_x_date(limits=as.Date(c("2012-01-01", "2012-12-31"))) + 
  ylab("") + xlab("") + # ggtitle("IgA by date of birth") + 
  theme_bw() # + theme(plot.title=element_text(hjust=0.5)) +

BPm_IgA = ggplot(data, aes(x=factor(month), y=IgATiter)) + 
  geom_dotplot(binaxis="y",stackdir='center',binwidth = 0.05, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) + 
  scale_y_log10(limits=c(10, 30000)) + 
  # scale_y_continuous(limits=c(10, 4000)) +
  scale_x_discrete(labels=month.abb) + 
  ylab("IgA (U/mL, log-scale)") + xlab("") + # ggtitle("IgA by month of birth") + 
  theme_bw() + # theme(plot.title=element_text(hjust=0.5)) + 
  stat_compare_means(method = "kruskal.test", label.y = 4.25, label.x = 2, 
                     label.sep=": ")

### IgG ----------------------------------------------------

BPd_IgG = ggplot(data, aes(x=DOB, y=IgGTiter)) + 
  geom_point() + 
  geom_smooth(method=gam, formula=y ~ s(x, bs="cc", k=12)) + 
  # scale_y_log10(limits=c(10, 12000)) + 
  ylab("") + xlab("") + # ggtitle("IgG by DOB") +
  scale_x_date(limits=as.Date(c("2012-01-01", "2012-12-31"))) + 
  theme_bw()

BPm_IgG = ggplot(data, aes(x=factor(month), y=IgGTiter)) + 
  geom_dotplot(binaxis="y",stackdir='center',binwidth = 0.05, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) + 
  # scale_y_log10(limits=c(10, 30000)) + 
  # scale_y_continuous(limits=c(10, 4000)) +
  scale_x_discrete(labels=month.abb) + 
  ylab("IgG (U/mL, log-scale)") + xlab("") + # ggtitle("IgG by MOB") +
  theme_bw() +
  stat_compare_means(method = "kruskal.test", label.y = 4.25, label.x = 2, 
                     label.sep=": ")

BPm_together = data %>%
  pivot_longer(cols = c(IgATiter, IgGTiter), 
               names_to = "antibody_type", 
               values_to = "titer") %>%
  mutate(antibody_type = factor(antibody_type, 
                               levels = c("IgATiter", "IgGTiter"),
                               labels = c("IgA", "IgG"))) %>% 
  ggplot(aes(x = factor(month), y = titer, fill = antibody_type, color=antibody_type)) +
  geom_dotplot(binaxis="y", stackdir='center', scale = "width",
               binwidth = 0.05, fill="gray50",
               stroke=0, position = position_dodge(width = 0.8),
               show.legend = F) +
  # geom_jitter(alpha = 0.6, size = 0.8,
  #           position = position_jitterdodge(dodge.width = 0.75)) + 
  geom_boxplot(position = position_dodge(width = 0.75), color = "gray50", width=0.7,
               outlier.shape = NA,  # Remove outliers as requested
               alpha = 0.7) +
  scale_fill_manual(values = c("IgA" = "red", "IgG" = "blue"),
                    name = "Antibody Type") +
  scale_color_manual(values = c("IgA" = "red", "IgG" = "blue"),
                    name = "Antibody Type") +
  scale_y_log10(labels = scales::comma) +
  # scale_y_continuous(limits=c(0, 4000), labels = comma,  expand = c(0, 0)) + 
  # coord_cartesian(ylim = c(0, 4000)) + 
  scale_x_discrete(labels = month.abb) +
  labs(x = "Month", 
       y = "Antibody concentration\n(U/mL, log scale)",
       title = "IgA and IgG Antibodies") +
       # subtitle = "Boxplots show median (GMT), IQR, and whiskers without outliers"
       # caption = "Kruskall-Wallis p-values for homogeneity: IgA = 0.080, IgG = 0.025") +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        panel.background = element_blank(),
        plot.background = element_blank()) 

# Display the plot
print(BPm_together)
ggsave("./out/inc_titers.pdf", height = 4.5, width = 6, units="in", dpi=300, bg=NULL)

### all desc graphs together ------


load("./out/ccf_gamma_uncertainty_lsd.Rdata")
monthly_incidence_lsd = monthly_incidence
load("./out/ccf_gamma_uncertainty_msd.Rdata")
monthly_incidence_msd = monthly_incidence
load("./out/ccf_gamma_uncertainty_total.Rdata")
monthly_incidence_total = monthly_incidence

# From the cross_corr_catalytics code. 
load("./out/ccf_data_uncertainty_lsd_plot.Rdata")
p_ccf_uncertainty_lsd = p_ccf_uncertainty
load("./out/ccf_data_uncertainty_msd_plot.Rdata")
p_ccf_uncertainty_msd = p_ccf_uncertainty
load("./out/ccf_data_uncertainty_total_plot.Rdata")
p_ccf_uncertainty_total = p_ccf_uncertainty

ggmonth1a_lsd = monthly_incidence_lsd[c(3:12, 1:2),] %>% # epi_desc_msd[4:15,] %>% 
  mutate(Group = as.Date(paste0(Group, "-01"), format="%Y-%m-%d")) %>% 
  mutate(month = as.numeric(format(Group, "%m"))) %>%
  mutate(month = factor(month, levels=1:12, labels=month.abb)) %>%
  mutate(lower = ifelse(lower<0.5, 0.5, lower)) %>% 
  mutate(mean = ifelse(mean<0.5, 0.5, mean)) %>% 
ggplot(aes(x=month, y=mean)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) + 
  # geom_vline(aes(xintercept=as.Date("2011-12-15"))) + 
  theme_bw() + 
  theme(panel.grid.minor=element_blank(),
        plot.margin = margin(t = 10, r = 1, b = 1, l = 1),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  scale_y_log10(limits=c(0.5, 60)) + 
  # scale_x_date(date_breaks = "1 month",
    # labels = function(x) format(x, "%b")) + 
  xlab(NULL) + ylab("Cases per 100 py") + 
  ggtitle("GEMS-1a, incidence of rotavirus-attributable LSD, children <5y") + 
  theme(plot.title = element_text(hjust=0.5, size=12))

ggmonth1a_msd = monthly_incidence_msd[c(3:12, 1:2),] %>% # epi_desc_msd[4:15,] %>% 
  mutate(Group = as.Date(paste0(Group, "-01"), format="%Y-%m-%d")) %>% 
  mutate(month = as.numeric(format(Group, "%m"))) %>%
  mutate(month = factor(month, levels=1:12, labels=month.abb)) %>%
  mutate(lower = ifelse(lower<0.5, 0.5, lower)) %>% 
  mutate(mean = ifelse(mean<0.5, 0.5, mean)) %>% 
ggplot(aes(x=month, y=mean)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) + 
  # geom_vline(aes(xintercept=as.Date("2011-12-15"))) + 
  theme_bw() + 
  theme(panel.grid.minor=element_blank(),
        plot.margin = margin(t = 10, r = 1, b = 1, l = 1),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  scale_y_log10(limits=c(0.5, 60)) + 
  # scale_x_date(date_breaks = "1 month",
    # labels = function(x) format(x, "%b")) + 
  xlab(NULL) + ylab("Cases per 100 py") + 
  ggtitle("GEMS-1a, incidence of rotavirus-attributable MSD, children <5y") + 
  theme(plot.title = element_text(hjust=0.5, size=12))

ggmonth1a_total = monthly_incidence_total[c(3:12, 1:2),] %>% # epi_desc_msd[4:15,] %>% 
  mutate(Group = as.Date(paste0(Group, "-01"), format="%Y-%m-%d")) %>% 
  mutate(month = as.numeric(format(Group, "%m"))) %>%
  mutate(month = factor(month, levels=1:12, labels=month.abb)) %>%
  mutate(lower = ifelse(lower<0.5, 0.5, lower)) %>% 
  mutate(mean = ifelse(mean<0.5, 0.5, mean)) %>% 
ggplot(aes(x=month, y=mean)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower, ymax = upper)) + 
  # geom_vline(aes(xintercept=as.Date("2011-12-15"))) + 
  theme_bw() + 
  theme(panel.grid.minor=element_blank(),
        plot.margin = margin(t = 10, r = 1, b = 1, l = 1),
        panel.background = element_blank(),
        plot.background = element_blank()) + 
  scale_y_log10(limits=c(0.5, 60)) + 
  # scale_y_continuous(limits=c(0, 60)) + 
  # scale_x_date(date_breaks = "1 month",
    # labels = function(x) format(x, "%b")) + 
  xlab(NULL) + ylab("Cases per 100 py") + 
  ggtitle("GEMS-1a, incidence of rotavirus-attributable diarrhea, children <5y") + 
  theme(plot.title = element_text(hjust=0.5, size=12))

# Display the plot
plot_grid(ggmonth1a_lsd, ggmonth1a_msd, 
          p_ccf_uncertainty_lsd, p_ccf_uncertainty_msd,
          labels = "AUTO", label_size = 12, 
          ncol=2, rel_heights = c(0.725, 1), align="v", axis="lr") 
ggsave("./out/titers_gems_1a_lsd_msd.png", height = 20, width = 30, units="cm", dpi=300, bg=NULL)

```
```{r main-ccf, fig.width=7, fig.height=11, message=F, warning=F}

plot_grid(BPm_together, ggmonth1a_total, p_ccf_uncertainty_total,
          labels = "AUTO", label_size = 12, 
          ncol=1, rel_heights = c(0.89, 0.725, 1), align="v", axis="rl")
ggsave("./out/titers_inc_ccf.png", height = 30, width = 17.5, units="cm", dpi=300, bg=NULL)

```

# ANOVA

```{r anova, fig.width=5, fig.height=7, warning=F, message=F}

## IgA ------
IgA_log.aov <- anova(lm(log(IgATiter) ~ factor(month, levels=c(10, 1:9, 11:12)), 
                        data = data))
#IgA_log.aov %>% as_flextable(show_coltype=F)

fit1A = lm(log(IgATiter) ~ factor(month, levels=c(10, 1:9, 11:12)), data = data)
# plot(fit1A, c(2,4))
# plot(log2(data$IgATiter), cooks.distance(fit1A))
# plot(residuals(fit1A), cooks.distance(fit1A))

names(fit1A$coefficients) = c("Intercept", month.name[1:9], month.name[11:12])

# fit1A %>% as_flextable(show_coltype=F)

tmp_coeff = summary(fit1A)$coefficients
tmp = tmp_coeff %>% as.data.frame() %>%
  mutate(est = exp(Estimate),
         lci = exp(Estimate-1.96*`Std. Error`), 
         uci = exp(Estimate+1.96*`Std. Error`)) %>% 
  mutate(est_ci = paste0(my_format_mod(est, 2, 3), " (", 
                         my_format_mod(lci, 2, 3), ", ", my_format_mod(uci, 2, 3), ")")) %>% 
  mutate(pstar = ifelse(`Pr(>|t|)`<0.001, "***", 
                        ifelse(`Pr(>|t|)`<0.01, "**", 
                               ifelse(`Pr(>|t|)`<0.05, "*", ""))))

fit1Apretty = data.frame(Month = c("Intercept", month.name[1:12]), 
                         Est_A = c(tmp$est_ci[1:10], "Ref", tmp$est_ci[11:12]),
                         P_A = c(tmp$pstar[1:10], "", tmp$pstar[11:12]))

tmp_1A_sum = summary(fit1A)
tmp_1A_sum_F = tmp_1A_sum$fstatistic["value"]
tmp_1A_sum_numdf = tmp_1A_sum$fstatistic["numdf"]
tmp_1A_sum_dendf = tmp_1A_sum$fstatistic["dendf"]

tmp_ft_A = data.frame(Month = c("N. obs", "SE Residuals", "R2; Adj-R2", "F-stat", "P-value"), 
                      Est_A = c(length(fit1A$fitted.values), 
                                my_format(sd(fit1A$residuals), 2),
                                paste0(my_format(summary(fit1A)$r.squared, 2), "; ", my_format(summary(fit1A)$adj.r.squared, 2)),
                                my_format(tmp_1A_sum_F, 2),
                                my_smallvalue(pf(tmp_1A_sum_F, tmp_1A_sum_numdf, tmp_1A_sum_dendf, lower.tail = F), 3)))


# tmp_ft_A = fit1A %>% as_flextable() %>%
#   delete_part(part="body") %>% 
#   delete_part(part="header") 

tmp_pretty_A = fit1Apretty %>% as_flextable(max_row = 13, show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header")

# vtmp_pretty_A$footer = tmp_ft_A$footer

# If you wanted to predict the means, here it is: 
fit1Apred = predict(fit1A, newdata=data.frame(month = factor(1:12)), interval="confidence")
# 2^(fit1Apred) %>% as_tibble %>% as_flextable(max_rows=12)
  
## IgG ------
IgG_log.aov = anova(lm(log(IgGTiter) ~ factor(month, levels=c(10, 1:9, 11:12)), data = data))
# lm(log2(IgGTiter) ~ factor(month, levels=c(10, 1:9, 11:12)), data = data) %>% summary
fit1G = lm(log(IgGTiter) ~ factor(month, levels=c(10, 1:9, 11:12)), data = data) # %>% summary

anova_IgG_resid_kurtosis = moments::kurtosis(fit1G$residuals)
anova_IgG_resid_skewness = moments::skewness(fit1G$residuals)
anova_IgG_resid_jarque_bera = moments::jarque.test(fit1G$residuals)

names(fit1G$coefficients) = c("Intercept", month.name[1:9], month.name[11:12])

tmp_coeff = summary(fit1G)$coefficients
tmp = tmp_coeff %>% as.data.frame() %>%
  mutate(est = exp(Estimate),
         lci = exp(Estimate-1.96*`Std. Error`), 
         uci = exp(Estimate+1.96*`Std. Error`)) %>% 
  mutate(est_ci = paste0(my_format_mod(est, 2, 3), " (", 
                my_format_mod(lci, 2, 3), ", ", my_format_mod(uci, 2, 3), ")")) %>% 
  mutate(pstar = ifelse(`Pr(>|t|)`<0.001, "***", 
                ifelse(`Pr(>|t|)`<0.01, "**", 
                ifelse(`Pr(>|t|)`<0.05, "*", ""))))

fit1Gpretty = data.frame(Month = c("Intercept", month.name[1:12]), 
          Est_G = c(tmp$est_ci[1:10], "Ref", tmp$est_ci[11:12]),
          P_G = c(tmp$pstar[1:10], "", tmp$pstar[11:12]))

tmp_1G_sum = summary(fit1G)
tmp_1G_sum_F = tmp_1G_sum$fstatistic["value"]
# tmp_1G_sum_numdf = tmp_1G_sum$fstatistic["numdf"]
# tmp_1G_sum_dendf = tmp_1G_sum$fstatistic["dendf"]

tmp_ft_G = data.frame(Month = c("N. obs", "SE Residuals", "R2; Adj-R2", "F-stat", "P-value"), 
                      Est_G = c(length(fit1G$fitted.values), 
                                my_format(sd(fit1G$residuals), 2),
                                paste0(my_smallvalue(summary(fit1G)$r.squared, 2), "; ", my_smallvalue(summary(fit1G)$adj.r.squared, 2)),
                                my_format(tmp_1G_sum_F, 2),
                                my_smallvalue(fpvalue(fit1G)$Fp, 3)))
# pf(tmp_1G_sum_F, tmp_1G_sum_numdf, tmp_1G_sum_dendf, lower.tail = F)

fit1Apretty %>% 
  bind_rows(tibble(tmp_ft_A)) %>%
  full_join(fit1Gpretty %>% bind_rows(tmp_ft_G)) %>%
  as_flextable(max_row = 30, show_coltype=F, spread_first_col = TRUE) %>% 
  delete_part(part = "footer") %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  bg(i=14:18, bg = "gray80", part = "body") %>% 
  hline(i=13, border=officer::fp_border(style = "solid", width = 2), part="body") %>% 
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(1, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("Month", "Estimate", "P-value", "Estimate", "P-value")) %>%
  hline(i=1, border=officer::fp_border(style = "solid", width = 2), part="header") %>% 
  align(align = "center", part = "header") %>% 
  line_spacing(space = 1, part = "all") %>% 
  add_footer_lines("Signif. codes: 0 ≤ '***' < 0.001 < '**' < 0.01 < '*' < 0.05") %>% 
  italic(part="footer", italic=T) # %>% 

```

\newpage
# Seasonal models

## Seasonal regression (parametric)
```{r seasonalols-calculations, warning=F, message=F}
# Fitting sine wave with log2 -------------------
## IgA
lm_IgA_sine   = lm(log(IgATiter)~1+cos(2*pi*dayofyear/366)+sin(2*pi*dayofyear/366), data)
lm_IgA_sine_m = lm(log(IgATiter)~1+cos(2*pi*month/12)+sin(2*pi*month/12), data)

## IgG 
lm_IgG_sine   = lm(log(IgGTiter)~1+cos(2*pi*dayofyear/366)+sin(2*pi*dayofyear/366), data)
lm_IgG_sine_m = lm(log(IgGTiter)~1+cos(2*pi*month/12)+sin(2*pi*month/12), data)

# Other useful quick code.
# plot(lm_log2_IgA_sine, which=2)

# lm_log2_IgG_sine %>% as_flextable %>%
#   bg(bg = "black", part = "header") %>% 
#   color(color = "white", part = "header") %>%
#   bg(bg = "gray80", part = "footer")

## Predictions -----

newpred_day   = data.frame(dayofyear=seq(1, 366))
newpred_month = data.frame(month=seq(1, 12))

sine_day_design = matrix(c(rep(1,length(newpred_day$dayofyear)), # intercept
                         cos(2*pi*newpred_day$dayofyear/366), # days of the year, cosine-d
                         sin(2*pi*newpred_day$dayofyear/366)), # days of the year, sine-d
                         ncol=3, byrow=F) # put this into 3 columns, fill the columns before you fill the rows

sine_month_design = matrix(c(rep(1,length(newpred_month$month)), # intercept
                         cos(2*pi*newpred_month$month/12), # days of the year, cosine-d
                         sin(2*pi*newpred_month$month/12)), # days of the year, sine-d
                         ncol=3, byrow=F) # put this into 3 columns, fill the columns before you fill the rows

# iterates of the coefficients
tmp_sim_IgA   = mvrnorm(1000, coef(lm_IgA_sine),   Sigma=vcov(lm_IgA_sine))
tmp_sim_IgA_m = mvrnorm(1000, coef(lm_IgA_sine_m), Sigma=vcov(lm_IgA_sine_m))
tmp_sim_IgG   = mvrnorm(1000, coef(lm_IgG_sine),   Sigma=vcov(lm_IgG_sine))
tmp_sim_IgG_m = mvrnorm(1000, coef(lm_IgG_sine_m), Sigma=vcov(lm_IgG_sine_m))

IgA_sine_pred_iter   = exp(sine_day_design %*% t(tmp_sim_IgA)) # t() bc the inner dimensions have to match (remember matrix multiplication)
IgA_sine_pred_iter_m = exp(sine_month_design %*% t(tmp_sim_IgA_m)) 
IgG_sine_pred_iter   = exp(sine_day_design %*% t(tmp_sim_IgG)) 
IgG_sine_pred_iter_m = exp(sine_month_design %*% t(tmp_sim_IgG_m)) 
# code for quick checks
# matplot(t(IgA_sine_pred_iter %>% apply(2, quantile, c(0.5, 0.025, 0.975))), type="l")

# Peaks and troughs - timing
IgA_sine_summ_d_min_ci  = IgA_sine_pred_iter %>% apply(2, which.min) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgA_sine_summ_d_min_iqr = IgA_sine_pred_iter %>% apply(2, which.min) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgA_sine_summ_d_max_ci  = IgA_sine_pred_iter %>% apply(2, which.max) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgA_sine_summ_d_max_iqr = IgA_sine_pred_iter %>% apply(2, which.max) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat

IgA_sine_summ_m_min_ci  = IgA_sine_pred_iter_m %>% apply(2, which.min) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgA_sine_summ_m_min_iqr = IgA_sine_pred_iter_m %>% apply(2, which.min) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgA_sine_summ_m_max_ci  = IgA_sine_pred_iter_m %>% apply(2, which.max) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgA_sine_summ_m_max_iqr = IgA_sine_pred_iter_m %>% apply(2, which.max) %>% med_iqr %>% month.abb[.] %>% time_range_concat

IgG_sine_summ_d_min_ci  = IgG_sine_pred_iter %>% apply(2, which.min) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgG_sine_summ_d_min_iqr = IgG_sine_pred_iter %>% apply(2, which.min) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgG_sine_summ_d_max_ci  = IgG_sine_pred_iter %>% apply(2, which.max) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgG_sine_summ_d_max_iqr = IgG_sine_pred_iter %>% apply(2, which.max) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat

IgG_sine_summ_m_min_ci  = IgG_sine_pred_iter_m %>% apply(2, which.min) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgG_sine_summ_m_min_iqr = IgG_sine_pred_iter_m %>% apply(2, which.min) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgG_sine_summ_m_max_ci  = IgG_sine_pred_iter_m %>% apply(2, which.max) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgG_sine_summ_m_max_iqr = IgG_sine_pred_iter_m %>% apply(2, which.max) %>% med_iqr %>% month.abb[.] %>% time_range_concat

# Peaks and troughs - antibody levels
IgAab_sine_summ_d_min_ci = IgA_sine_pred_iter %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgAab_sine_summ_d_max_ci = IgA_sine_pred_iter %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)
IgAab_sine_summ_m_min_ci = IgA_sine_pred_iter_m %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgAab_sine_summ_m_max_ci = IgA_sine_pred_iter_m %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)

IgGab_sine_summ_d_min_ci = IgG_sine_pred_iter %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgGab_sine_summ_d_max_ci = IgG_sine_pred_iter %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)
IgGab_sine_summ_m_min_ci = IgG_sine_pred_iter_m %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgGab_sine_summ_m_max_ci = IgG_sine_pred_iter_m %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)

# amplitude - absolute
IgAab_sine_summ_d_amp_ci = IgA_sine_pred_iter %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgAab_sine_summ_m_amp_ci = IgA_sine_pred_iter_m %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgGab_sine_summ_d_amp_ci = IgG_sine_pred_iter %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgGab_sine_summ_m_amp_ci = IgG_sine_pred_iter_m %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)

# amplitude - ratio
myratio = function(x){x[2]/x[1]}
IgAab_sine_summ_d_rat_ci = IgA_sine_pred_iter %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgAab_sine_summ_m_rat_ci = IgA_sine_pred_iter_m %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgGab_sine_summ_d_rat_ci = IgG_sine_pred_iter %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgGab_sine_summ_m_rat_ci = IgG_sine_pred_iter_m %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)

sine_models = list("IgA, DOB" = lm_IgA_sine, "IgA, MOB" = lm_IgA_sine_m, "IgG, DOB" = lm_IgG_sine, "IgG, MOB" = lm_IgG_sine_m)

fstats = sapply(1:length(sine_models), function(i){fpvalue(sine_models[[i]])}) %>% unlist 
fstats = data.frame(part="gof", term="P-value", 
                    model_name = names(sine_models), 
                    Fp = fstats) %>% 
  pivot_wider(names_from = model_name, values_from = Fp) %>% 
  mutate(across(names(sine_models), ~ my_smallvalue(.x, 3)))

# peaks and troughs - time
lm_peak_ci = tibble("IgA, DOB" = IgA_sine_summ_d_max_ci, "IgA, MOB" = IgA_sine_summ_m_max_ci, 
               "IgG, DOB" = IgG_sine_summ_d_max_ci, "IgG, MOB" = IgG_sine_summ_m_max_ci) %>% 
                  mutate(term = "Peak time")
lm_trough_ci = tibble("IgA, DOB" = IgA_sine_summ_d_min_ci, "IgA, MOB" = IgA_sine_summ_m_min_ci, 
                 "IgG, DOB" = IgG_sine_summ_d_min_ci, "IgG, MOB" = IgG_sine_summ_m_min_ci) %>% 
                  mutate(term = "Trough time")

# peaks and troughs - amplitude and ratio
lm_amp = tibble("IgA, DOB" = IgAab_sine_summ_d_amp_ci, "IgA, MOB" = IgAab_sine_summ_m_amp_ci, 
                "IgG, DOB" = IgGab_sine_summ_d_amp_ci, "IgG, MOB" = IgGab_sine_summ_m_amp_ci) %>% 
                    mutate(term = "Amplitude") %>% relocate(term)
lm_rat = tibble("IgA, DOB" = IgAab_sine_summ_d_rat_ci, "IgA, MOB" = IgAab_sine_summ_m_rat_ci, 
                "IgG, DOB" = IgGab_sine_summ_d_rat_ci, "IgG, MOB" = IgGab_sine_summ_m_rat_ci) %>% 
                    mutate(term = "Peak:Trough ratio") %>% relocate(term)

# peaks and troughs - antibodies
lm_peak_ab_ci = tibble("IgA, DOB" = IgAab_sine_summ_d_max_ci, "IgA, MOB" = IgAab_sine_summ_m_max_ci, 
                    "IgG, DOB" = IgGab_sine_summ_d_max_ci, "IgG, MOB" = IgGab_sine_summ_m_max_ci) %>% 
                    mutate(term = "Peak antibodies") %>% relocate(term)
lm_trough_ab_ci = tibble("IgA, DOB" = IgAab_sine_summ_d_min_ci, "IgA, MOB" = IgAab_sine_summ_m_min_ci, 
                    "IgG, DOB" = IgGab_sine_summ_d_min_ci, "IgG, MOB" = IgGab_sine_summ_m_min_ci) %>% 
                    mutate(term = "Trough antibodies") %>% relocate(term)

ms = modelsummary(sine_models,
             fmt=NULL,
             estimate  = "{my_format_mod(exp(estimate),2,10)} ({my_format_mod(exp(conf.low),2,10)}, {my_format_mod(exp(conf.high),2,10)}){stars}",
             coef_rename = c("(Intercept)" = "Intercept",
                             "cos(2 * pi * dayofyear/366)" = "Cosine", 
                             "sin(2 * pi * dayofyear/366)" = "Sine",
                             "cos(2 * pi * month/12)" = "Cosine", 
                             "sin(2 * pi * month/12)" = "Sine"),
             gof_map = c("nobs", "r.squared", "adj.r.squared", "F"),
             statistic = NULL,
             output="data.frame") 

combined_summary = bind_rows(ms, fstats) # lm_amp, lm_rat, lm_peak_ci, lm_trough_ci,
# combined_summary = combined_summary[c(1:3, 4:7, 12),] # c(1:3, 8:11, 4:7, 12)

```

## Splines (non-parametric)
```{r seasonalsplines-calculations, warning=F, message=F}
## GAMs ----------------------------------------------------

### GAMs by day 
IgA_spline_k = fit_gam_model_kout(data, "IgATiter", "dayofyear") %>% which.min() # the k for the model below. 
IgA_spline = gam(log(IgATiter)~s(dayofyear, bs='cc', k=IgA_spline_k), family=gaussian, data = data)
# plot(IgA_spline, scale = 0)

IgG_spline_k = fit_gam_model_kout(data, "IgGTiter", "dayofyear") %>% which.min() # the k for the model below.
# The difference is minimal, so we will take the minimum k

IgG_spline = gam(log(IgGTiter)~s(dayofyear, bs="cc", k=3), family=gaussian, data = data)
# plot(IgG_spline, scale = 0)

### GAMs by month (robustness check) 
IgA_spline_m_k = fit_gam_model_kout(data, "IgATiter", "month") %>% which.min() # the k for the model below.
IgA_spline_m = gam(log(IgATiter)~s(month,bs="cc", k=IgA_spline_m_k), family=gaussian, data = data)
# plot(IgA_spline_m, scale = 0)

IgG_spline_m_k = fit_gam_model_kout(data, "IgGTiter", "month") %>% which.min() # the k for the model below. 
# The difference is minimal, so we will take the minimum k

IgG_spline_m = gam(log(IgGTiter)~s(month,bs="cc", k=IgG_spline_m_k), family=gaussian, data = data)
# plot(IgG_spline_m, scale = 0)

# Some basic code to check gam models
# coef(IgA_spline)
# gam.check(IgA_spline) # this shows that log is necessary. When I did the linear one, the Q-Q plot was a mess.
# summary(IgA_spline)
# plot(IgA_spline, scale = 0)
# anova(IgA_spline)

# Flextable to summarize one model alone (almost an ANOVA table for GAMs)
# as_flextable(IgA_spline) %>%
#   bg(bg = "black", part = "header") %>% 
#   color(color = "white", part = "header") %>%
#   bg(i=2, bg = "black", part = "body") %>% 
#   color(i=2, color = "white", part = "body") %>%
#   bg(bg = "gray80", part = "footer")

### Predictions and summaries
newpred_day = data.frame(dayofyear=seq(1, 366))
newpred_month = data.frame(month=seq(1, 12))

# Getting iterates out of the models -----
IgA_spline_pred   = gam_predictions(IgA_spline, newpred_day, 50000) %>% exp()
IgA_spline_m_pred = gam_predictions(IgA_spline_m, newpred_month, 50000) %>% exp()
IgG_spline_pred   = gam_predictions(IgG_spline, newpred_day, 50000) %>% exp()
IgG_spline_m_pred = gam_predictions(IgG_spline_m, newpred_month, 50000) %>% exp()

# IgG_spline_resp = predict.gam(IgG_spline, newdata=newpred_day, type="response", se.fit=T) 

# Summarizing the peaks and the troughs of each model with 95% CIs and Median [IQR]
# IgA_spline_summ_d_min_iqr = IgA_spline_pred %>% apply(2, which.min) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgA_spline_summ_d_min_ci = IgA_spline_pred %>% apply(2, which.min) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgA_spline_summ_d_max_iqr = IgA_spline_pred %>% apply(2, which.max) %>% med_iqr %>% 
#                               +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgA_spline_summ_d_max_ci = IgA_spline_pred %>% apply(2, which.max) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat

# IgA_spline_summ_m_min_iqr = IgA_spline_m_pred %>% apply(2, which.min) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgA_spline_summ_m_min_ci = IgA_spline_m_pred %>% apply(2, which.min) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgA_spline_summ_m_max_iqr = IgA_spline_m_pred %>% apply(2, which.max) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgA_spline_summ_m_max_ci = IgA_spline_m_pred %>% apply(2, which.max) %>% mean_ci %>% month.abb[.] %>% time_range_concat

# IgG_spline_summ_d_min_iqr = IgG_spline_pred %>% apply(2, which.min) %>% med_iqr  %>% 
                              #+as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgG_spline_summ_d_min_ci = IgG_spline_pred %>% apply(2, which.min) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
# IgG_spline_summ_d_max_iqr = IgG_spline_pred %>% apply(2, which.max) %>% med_iqr %>% 
                              #+as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat
IgG_spline_summ_d_max_ci = IgG_spline_pred %>% apply(2, which.max) %>% mean_ci %>% 
                              +as.Date("2012-01-01") %>% format("%b %d") %>% time_range_concat

# IgG_spline_summ_m_min_iqr = IgG_spline_m_pred %>% apply(2, which.min) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgG_spline_summ_m_min_ci = IgG_spline_m_pred %>% apply(2, which.min) %>% mean_ci %>% month.abb[.] %>% time_range_concat
# IgG_spline_summ_m_max_iqr = IgG_spline_m_pred %>% apply(2, which.max) %>% med_iqr %>% month.abb[.] %>% time_range_concat
IgG_spline_summ_m_max_ci = IgG_spline_m_pred %>% apply(2, which.max) %>% mean_ci %>% month.abb[.] %>% time_range_concat

# antibody levels at peak and trough
IgAab_spline_summ_d_min_ci = IgA_spline_pred %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgAab_spline_summ_d_max_ci = IgA_spline_pred %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)
IgAab_spline_summ_m_min_ci = IgA_spline_m_pred %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgAab_spline_summ_m_max_ci = IgA_spline_m_pred %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)

IgGab_spline_summ_d_min_ci = IgG_spline_pred %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgGab_spline_summ_d_max_ci = IgG_spline_pred %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)
IgGab_spline_summ_m_min_ci = IgG_spline_m_pred %>% apply(2, min) %>% mean_ci %>% ci_string_dec(0)
IgGab_spline_summ_m_max_ci = IgG_spline_m_pred %>% apply(2, max) %>% mean_ci %>% ci_string_dec(0)

# amplitude - absolute
IgAab_spline_summ_d_amp_ci = IgA_spline_pred %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgAab_spline_summ_m_amp_ci = IgA_spline_m_pred %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgGab_spline_summ_d_amp_ci = IgG_spline_pred %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)
IgGab_spline_summ_m_amp_ci = IgG_spline_m_pred %>% apply(2, range) %>% apply(2, diff) %>% mean_ci %>% ci_string_dec(0)

# amplitude - ratio
myratio = function(x){x[2]/x[1]}
IgAab_spline_summ_d_rat_ci = IgA_spline_pred %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgAab_spline_summ_m_rat_ci = IgA_spline_m_pred %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgGab_spline_summ_d_rat_ci = IgG_spline_pred %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)
IgGab_spline_summ_m_rat_ci = IgG_spline_m_pred %>% apply(2, range) %>% apply(2, myratio) %>% mean_ci %>% ci_string_dec(2)

# Summaries -----
## antibodies
gam_peak_ab_ci = tibble("IgA, DOB" = IgAab_spline_summ_d_max_ci, "IgA, MOB" = IgAab_spline_summ_m_max_ci, 
                    "IgG, DOB" = IgGab_spline_summ_d_max_ci, "IgG, MOB" = IgGab_spline_summ_m_max_ci) %>% 
                    mutate(term = "Peak antibodies") %>% relocate(term)
gam_trough_ab_ci = tibble("IgA, DOB" = IgAab_spline_summ_d_min_ci, "IgA, MOB" = IgAab_spline_summ_m_min_ci, 
                    "IgG, DOB" = IgGab_spline_summ_d_min_ci, "IgG, MOB" = IgGab_spline_summ_m_min_ci) %>% 
                    mutate(term = "Trough antibodies") %>% relocate(term)
## time
gam_peak_ci = tibble("IgA, DOB" = IgA_spline_summ_d_max_ci, "IgA, MOB" = IgA_spline_summ_m_max_ci, 
                "IgG, DOB" = IgG_spline_summ_d_max_ci, "IgG, MOB" = IgG_spline_summ_m_max_ci) %>% 
                    mutate(term = "Peak time") %>% relocate(term)
gam_trough_ci = tibble("IgA, DOB" = IgA_spline_summ_d_min_ci, "IgA, MOB" = IgA_spline_summ_m_min_ci, 
                  "IgG, DOB" = IgG_spline_summ_d_min_ci, "IgG, MOB" = IgG_spline_summ_m_min_ci) %>% 
                    mutate(term = "Trough time") %>% relocate(term)
## amplitude and ratio
gam_amp = tibble("IgA, DOB" = IgAab_spline_summ_d_amp_ci, "IgA, MOB" = IgAab_spline_summ_m_amp_ci, 
            "IgG, DOB" = IgGab_spline_summ_d_amp_ci, "IgG, MOB" = IgGab_spline_summ_m_amp_ci) %>% 
                    mutate(term = "Amplitude") %>% relocate(term)
gam_rat = tibble("IgA, DOB" = IgAab_spline_summ_d_rat_ci, "IgA, MOB" = IgAab_spline_summ_m_rat_ci, 
            "IgG, DOB" = IgGab_spline_summ_d_rat_ci, "IgG, MOB" = IgGab_spline_summ_m_rat_ci) %>% 
                    mutate(term = "Peak:Trough ratio") %>% relocate(term)

## Pulling GAM models together -----------------------
gam_model_intercepts = c("IgA, DOB" = (extract_Int_CI(IgA_spline)   %>% exp() %>% ci_string_dec(0)), 
                         "IgA, MOB" = (extract_Int_CI(IgA_spline_m) %>% exp() %>% ci_string_dec(0)),
                         "IgG, DOB" = (extract_Int_CI(IgG_spline)   %>% exp() %>% ci_string_dec(0)),
                         "IgG, MOB" = (extract_Int_CI(IgG_spline_m) %>% exp() %>% ci_string_dec(0))) %>% 
                           tibble::enframe(name = "model", value = "Intercept") 

gam_k = c("IgA, DOB" = IgA_spline_k, "IgA, MOB" = IgA_spline_m_k, 
          "IgG, DOB" = IgG_spline_k, "IgG, MOB" = IgG_spline_m_k) %>% 
             tibble::enframe(name = "model", value = "k") 

all_tidy_gams = bind_rows( # modelsummary doesn't play nice with gam objects, hence tidy()
  mutate(tidy(IgA_spline,   type = "smooth"), model = "IgA, DOB"),
  mutate(tidy(IgA_spline_m, type = "smooth"), model = "IgA, MOB"),
  mutate(tidy(IgG_spline,   type = "smooth"), model = "IgG, DOB"),
  mutate(tidy(IgG_spline_m, type = "smooth"), model = "IgG, MOB")) %>% 
  full_join(gam_model_intercepts) %>% 
  # full_join(gam_amp) %>% 
  # full_join(gam_rat) %>% 
  # full_join(gam_peak_ci) %>% 
  # full_join(gam_trough_ci) %>% 
  full_join(gam_k) %>% 
  relocate(model, term, Intercept, edf, k) %>% # Amplitude, PeakTroughRatio, peak_CI, trough_CI,
  mutate(across(c("edf", "statistic"), ~ my_format(.x, 2))) %>% 
  mutate(p.value = my_smallvalue(p.value, 3))
  
all_tidy_gams_df = all_tidy_gams %>% t() %>% as.data.frame()
colnames(all_tidy_gams_df) = all_tidy_gams_df[1,]
all_tidy_gams_df = all_tidy_gams_df[-c(1:2),]
all_tidy_gams_df$prop = c("Intercept", 
                          #"Amplitude", "Peak:Through Ratio",
                          #"Peak time", "Trough time", 
                          "EDF", "Basis dimension (k)", "Ref DF", "F", "P-value")

```

<!-- ## Putting both types of models together -->
```{r allmodelstable, warning=F, message=F, fig.height=5, fig.width=5}

# sine wave
all_tidy_sine_df_main = combined_summary %>% 
  dplyr::select(-part, -statistic) %>% 
  bind_rows(lm_peak_ab_ci) %>%
  bind_rows(lm_trough_ab_ci) %>%
  bind_rows(lm_amp) %>%
  bind_rows(lm_rat) %>%
  bind_rows(lm_peak_ci) %>%
  bind_rows(lm_trough_ci) %>%
  filter(term %in% c("Peak time", "Trough time", "Peak:Trough ratio", # "Amplitude", 
                     "Intercept", "Peak antibodies", "Trough antibodies",
                      "P-value")) %>%
  arrange(match(term, c("Peak time", "Trough time", "Peak:Trough ratio", # "Amplitude", 
                        "Intercept", "Peak antibodies", "Trough antibodies", 
                        "P-value"))) %>% 
  mutate(Models = "Par") %>% 
  mutate(across(where(is.character), ~ str_replace_all(., "\\*", ""))) %>%
  relocate(Models) %>% 
  mutate(term = c("Peak time", "Trough time", "Peak:Trough ratio", # "Amplitude", 
                  "Average antibodies", "Peak antibodies", "Trough antibodies", 
                  "P-value"))

# gams
all_tidy_gams_df_main = all_tidy_gams %>% 
  dplyr::select(model, Intercept, p.value) %>% 
  t() %>% as.data.frame()
colnames(all_tidy_gams_df_main) = all_tidy_gams_df_main[1,]
all_tidy_gams_df_main = all_tidy_gams_df_main[-1,]
all_tidy_gams_df_main$term = c("Average antibodies", "P-value")

all_tidy_gams_df_main = gam_peak_ci %>%
  bind_rows(gam_trough_ci) %>%
  bind_rows(gam_rat) %>%
  bind_rows(all_tidy_gams_df_main) %>%
  bind_rows(gam_peak_ab_ci) %>%
  bind_rows(gam_trough_ab_ci) %>%
  # bind_rows(gam_amp) %>%
  mutate(Models = "NonPar") %>% 
  relocate(Models, term) %>% 
  arrange(match(term, c("Peak time", "Trough time", "Peak:Trough ratio",
                        "Average antibodies", "Peak antibodies", "Trough antibodies", 
                        "P-value"))) 

# Combined # add amplitude and the ratios here... Consider if to remoce the others?
bind_rows(all_tidy_sine_df_main, all_tidy_gams_df_main) %>% 
  rename("IgA_DOB" = "IgA, DOB", "IgG_DOB" = "IgG, DOB", "IgA_MOB" = "IgA, MOB", "IgG_MOB" = "IgG, MOB") %>% 
  pivot_longer(cols=IgA_DOB:IgG_MOB) %>% 
  separate(name, into = c("Antibody", "Timing"), sep = "_") %>% 
  pivot_wider(names_from = c(Models, Timing), names_sep = "_", values_from = value) %>% 
  arrange(Antibody) %>% 
  as_grouped_data(groups="Antibody") %>% 
  # mutate(Models = factor(Models)) %>% 
  # as_grouped_data(groups = "Models") %>% 
  as_flextable() %>%
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>%
  color(color = "white", part = "header") %>%
  bg(i=c(8, 16), bg = "gray80", part = "body") %>% 
  bold(i=c(8, 16), bold=T, part="body") %>%
  italic(i=c(8, 16), italic=T, part="body") %>%
  bg(i=c(1, 9), bg = "gray30", part = "body") %>% 
  color(i=c(1, 9), color = "white", part = "body") %>%
  italic(i = c(1, 9), part = "body") %>% 
  compose(i = 1, j = 1, value = as_paragraph("IgA")) %>%
  compose(i = 9, j = 1, value = as_paragraph("IgG")) %>%
  hline(i=c(4,12), border=officer::fp_border(style = "solid", color = "gray30", width = 1.5), part="body") %>% 
  hline(i=c(1,8,9), border=officer::fp_border(style = "solid", color = "gray30", width = 1.5), part="body") %>% 
  add_header_row(values = c("", "Parametric model", "Non-parametric model"), colwidths = c(1, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("", "Days", "Months", "Days", "Months")) %>%
  align(align = "center", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header") %>% 
  # line_spacing(space = 1, part = "all") %>% 
  fontsize(size = 11, part = "all") %>% 
  set_table_properties(layout='autofit', width=1,
                       word_description="Seasonal models. The intercept indicates the average antibody levels throughout the year, not the peaks or the troughs. Peak and trough values are means (95% CI). P-values come from the F-test on whether the seasonal curves fit the data significantly better than a null model with no seasonal patterns. P-values<0.05 indicate significance.")

dates_15th = seq.Date(from = as.Date("2012-01-15"), to = as.Date("2012-12-15"), by = "month")
month_midpoints = as.numeric(dates_15th - as.Date("2012-01-01") + 1)

all_datsum_list = list()
all_mods_list = list()

all_datsum_list[[1]] = IgA_data_month = data %>% group_by(month) %>% 
  summarize(mean = GMT(IgATiter), lci = GMT_CI_se_low(IgATiter), uci=GMT_CI_se_high(IgATiter)) %>% 
  mutate(month=month.abb[month], Ig = "IgA")
all_datsum_list[[2]] = IgG_data_month = data %>% group_by(month) %>% 
  summarize(mean = GMT(IgGTiter), lci = GMT_CI_se_low(IgGTiter), uci=GMT_CI_se_high(IgGTiter)) %>% 
  mutate(month=month.abb[month], Ig = "IgG")
  
get_pred_month_summary = function(pred_array){
  tmp = pred_array %>% apply(1, mean_ci) %>% as.data.frame %>% tibble
  colnames(tmp) = month.abb
  tmp$measure = c("mean", "lci", "uci")
  tmp = tmp %>% pivot_longer(cols=Jan:Dec, names_to = "month") %>% 
    pivot_wider(id_cols=month, names_from = measure, values_from = value)
  return(tmp)
}
  
all_mods_list[[1]] = IgA_sine_pred_iter[month_midpoints,] %>% get_pred_month_summary %>% 
        mutate(Ig = "IgA", Time = "Day", Model = "Parametric model")
all_mods_list[[2]] = IgA_sine_pred_iter_m %>% get_pred_month_summary %>% 
        mutate(Ig = "IgA", Time = "Month", Model = "Parametric model")
all_mods_list[[3]] = IgG_sine_pred_iter[month_midpoints,] %>% get_pred_month_summary %>% 
        mutate(Ig = "IgG", Time = "Day", Model = "Parametric model")
all_mods_list[[4]] = IgG_sine_pred_iter_m %>% get_pred_month_summary %>% 
        mutate(Ig = "IgG", Time = "Month", Model = "Parametric model")

all_mods_list[[5]] = IgA_spline_pred[month_midpoints,] %>% get_pred_month_summary %>% 
        mutate(Ig = "IgA", Time = "Day", Model = "Non-parametric model")
all_mods_list[[6]] = IgA_spline_m_pred %>% get_pred_month_summary %>% 
        mutate(Ig = "IgA", Time = "Month", Model = "Non-parametric model")
all_mods_list[[7]] = IgG_spline_pred[month_midpoints,] %>% get_pred_month_summary %>% 
        mutate(Ig = "IgG", Time = "Day", Model = "Non-parametric model")
all_mods_list[[8]] = IgG_spline_m_pred %>% get_pred_month_summary %>% 
        mutate(Ig = "IgG", Time = "Month", Model = "Non-parametric model")

all_datsum = bind_rows(all_datsum_list) %>% rename(Month = month) %>% 
  mutate(Month = base::as.Date(paste("15", Month, "2012"), format = "%d %b %Y"))
all_mods = bind_rows(all_mods_list) %>% rename(Month = month) %>% 
  mutate(Month = as.Date(paste("15", Month, "2012"), format = "%d %b %Y")) %>% 
  mutate(Models = factor(Model)) %>%
  mutate(Models = factor(Model, levels = c("Parametric model", "Non-parametric model")))

p_bothmodels = ggplot(all_mods, aes(x=Month, y=mean)) + 
  geom_ribbon(aes(ymin = lci, ymax = uci, fill=Models), alpha = 0.5) + 
  geom_line(aes(color=Models), show.legend = F) +
  facet_grid(Ig~Time) + 
  geom_point(data = all_datsum, aes(x=Month, y=mean)) + 
  geom_errorbar(data = all_datsum, aes(ymin=lci, ymax=uci)) + 
  theme_bw() + 
  theme(strip.background = element_blank()) + 
  theme(legend.position = "bottom", 
        legend.title = element_blank()) + 
  scale_x_date(date_labels = "%b", date_breaks = "3 month") + 
  scale_y_log10(limits=c(200, 2000)) + 
  ylab("Ig (U/mL, log-scale)") + xlab(NULL) +
  labs(caption = "Points and error bars: GMT (95% CI)")

p_bothmodels

ggsave("./out/seasonal_models.png", height = 15, width = 20, units="cm", dpi=300, bg=NULL)

```

\newpage
# Calculating impact 
```{r ZambiatoMali, warning=F}

# 1-OR^(log(peaktroughratio))*100

# Parametric:
# IgA_sine_pred_iter %>% apply(2, range)
# IgG_sine_pred_iter %>% apply(2, range)

myratio = function(x){x[2]/x[1]}

peak_trough_ratio_IgA = IgA_spline_pred %>% apply(2, range) %>% apply(2, myratio)
peak_trough_ratio_IgG = IgG_spline_pred %>% apply(2, range) %>% apply(2, myratio)

logmean = log(0.8)
logsd = (log(0.94) - log(0.68))/(2*1.96)

OR_zambia_sero_unadj = rlnorm(length(peak_trough_ratio_IgA), 
                            logmean, logsd)

logmean = log(0.78)
logsd = (log(0.92) - log(0.66))/(2*1.96)

OR_zambia_sero_adj = rlnorm(length(peak_trough_ratio_IgA), 
                            logmean, logsd)

# 0.80 (0.68-0.94) unadjusted and 0.78 (0.66-0.92) after adjustment for seasonality in Zambia. They only measured IgA.

# Moon et al: 
logmean = log(0.8)
logsd = (log(0.96) - log(0.66))/(2*1.96)

OR_southafrica = rlnorm(length(peak_trough_ratio_IgA), 
                            logmean, logsd)

logmean = log(0.72)
logsd = (log(0.94) - log(0.55))/(2*1.96)

OR_malawi_sero = rlnorm(length(peak_trough_ratio_IgA), 
                            logmean, logsd)
# Mandolo et al 2025 (Malawi): 0.72 (0.55-0.94)

print("CONTINUOUS SEASON ANALYSIS")
sim_sero_IgA_unadj = (1-OR_zambia_sero_unadj^(log2(peak_trough_ratio_IgA)))*100
print("IgA ratio from spline and sero-positivity, Chilengi effect")
sim_sero_IgA_unadj %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgG_unadj = (1-OR_zambia_sero_unadj^(log2(peak_trough_ratio_IgG)))*100
print("IgG ratio from spline and sero-positivity, Chilengi effect")
sim_sero_IgG_unadj %>% mean_ci %>% ci_string_dec(2)

sim_sero_IgA_mwi = (1-OR_southafrica^(log2(peak_trough_ratio_IgA)))*100
print("IgA ratio from spline and sero-positivity, Moon effect")
sim_sero_IgA_mwi %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgG_mwi = (1-OR_southafrica^(log2(peak_trough_ratio_IgG)))*100
print("IgG ratio from spline and sero-positivity, Moon effect")
sim_sero_IgG_mwi %>% mean_ci %>% ci_string_dec(2)

sim_sero_IgA_mwi = (1-OR_malawi_sero^(log2(peak_trough_ratio_IgA)))*100
print("IgA ratio from spline and sero-positivity, Mandolo effect")
sim_sero_IgA_mwi %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgG_mwi = (1-OR_malawi_sero^(log2(peak_trough_ratio_IgG)))*100
print("IgG ratio from spline and sero-positivity, Mandolo effect")
sim_sero_IgG_mwi %>% mean_ci %>% ci_string_dec(2)

print("BINARY SEASON ANALYSIS")
print("IgA ratio from two-season analysis and sero-positivity, Chilengi effect")
high_low_season_IgA_mean = log10(1.54)
high_low_season_IgA_sd = (log10(2.04) - log10(1.15))/(2*1.96)

high_low_season_IgA = 10^rnorm(length(peak_trough_ratio_IgA),
                        high_low_season_IgA_mean, high_low_season_IgA_sd)

sim_sero_IgA_hl_unadj = (1-OR_zambia_sero_unadj^log2(high_low_season_IgA))*100
sim_sero_IgA_hl_unadj %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgA_hl_zaf = (1-OR_southafrica^log2(high_low_season_IgA))*100
sim_sero_IgA_hl_zaf %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgA_hl_mwi = (1-OR_malawi_sero^log2(high_low_season_IgA))*100
sim_sero_IgA_hl_mwi %>% mean_ci %>% ci_string_dec(2)

print("IgG ratio from two-season analysis and sero-positivity")
high_low_season_IgG_mean = log10(1.39)
high_low_season_IgG_sd = (log10(1.85) - log10(1.04))/(2*1.96)

high_low_season_IgG = 10^rnorm(length(peak_trough_ratio_IgG),
                        high_low_season_IgG_mean, high_low_season_IgG_sd)

sim_sero_IgG_hl_unadj = (1-OR_zambia_sero_unadj^log2(high_low_season_IgG))*100
sim_sero_IgG_hl_unadj %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgG_hl_zaf = (1-OR_southafrica^log2(high_low_season_IgG))*100
sim_sero_IgG_hl_zaf %>% mean_ci %>% ci_string_dec(2)
sim_sero_IgG_hl_mwi = (1-OR_malawi_sero^log2(high_low_season_IgG))*100
sim_sero_IgG_hl_mwi %>% mean_ci %>% ci_string_dec(2)

```

\newpage 
# Appendix 

## Omitting IgA <7.5

```{r omitlowIgAappendix, warning=F, fig.width=7, fig.height=5, fig.cap="Correlation between IgA and IgG concentrations in mothers. The data points on the top left show us the samples for which IgA has degraded. Those data points have been omitted."}
tmp = fisher.test(as.matrix(IgAtoolow[,2:3]))

IgAtoolow %>%
  bind_rows(IgAtoolow_all) %>%
  rename(Month = month, `Under 7.5`= u7, `Over 7.5`= o7) %>% 
  as_flextable(max_row = 13, show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  bg(i=13, bg = "gray80", part = "body") %>% 
  add_footer_row(values = paste0("Fisher's Exact Test, p=", 
                                 format(round(tmp$p.value, 2), nsmall=2)),
                 colwidths = 3)

data_original %>% filter(!is.na(IgATiter), !is.na(IgGTiter)) %>%
  ggplot(aes(x=IgATiter, y=IgGTiter)) + 
    geom_point(aes(color = (IgATiter>7.50))) + 
    theme_bw() + theme(legend.position="bottom") + 
    scale_y_log10(limits=c(5, 10000), labels=scales::comma) + 
    scale_x_log10(limits=c(5, 10000), labels=scales::comma) + 
    coord_equal() + 
    ylab("IgG (U/mL, log-scale)") + xlab("IgA (U/mL, log-scale)") + 
    scale_color_manual(values = c("TRUE" = "black", "FALSE" = "darkred"), 
                     labels = c("TRUE" = "Included", "FALSE" = "Excluded")) +
  labs(color = "Status") + 
  stat_cor(method="spearman", label.x.npc=0.2, label.y.npc=0.175, label.sep="; ", p.accuracy=0.001) + 
  stat_cor(method="spearman", data = data, label.x.npc=0.2, label.y.npc=0.1, label.sep="; ", p.accuracy=0.001) +
  annotate("text", x = 400, y = 18.5, label = "(all data)", size = 4, hjust = 0) +
  annotate("text", x = 400, y = 10.5, label = "(data included only)", size = 4, hjust = 0)
# the one without the IgA<7.5

ggsave("./out/correlation_IgA_IgG.png", height = 6, width = 5.5, units="in", dpi=300, bg=NULL)

```

\newpage
## Seasonal regression (parametric)
This is a re-parameterization of time so that a cyclic, seasonal model can be run by ordinary OLS regression.

```{r seasonalols, fig.width=5, fig.height=5, warning=F, message=F}

combined_summary %>% dplyr::select(-part, -statistic) %>%
  as_flextable(show_coltype=F, max_row=12) %>% 
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>%
  color(color = "white", part = "header") %>%
  bg(i=4:8, bg = "gray80", part = "body") %>% # 8:12
  bold(i=8, bold=T, part="body") %>%
  italic(i=8, italic=T, part="body") %>%
  hline(i=3, border=officer::fp_border(style = "solid", width = 1.5), part="body") %>% 
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(1, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("", "Days", "Months", "Days", "Months")) %>%
  align(align = "center", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header") %>% 
  set_table_properties(layout='autofit', width=1,
                       word_description="Sine model regressions. The p-values are calculated from an F-test.")

fitsine_Ad_qq = gg_qq(residuals(lm_IgA_sine))
fitsine_Gd_qq = gg_qq(residuals(lm_IgG_sine))
fitsine_Am_qq = gg_qq(residuals(lm_IgA_sine_m))
fitsine_Gm_qq = gg_qq(residuals(lm_IgG_sine_m))

plot_grid(fitsine_Ad_qq, fitsine_Gd_qq, fitsine_Am_qq, fitsine_Gm_qq, ncol=2, nrow=2, labels="AUTO") 
```

\newpage
## Splines
```{r seasonalspline, fig.width=5, fig.height=5, warning=F, message=F}

# Final flextable
all_tidy_gams_df %>% 
  relocate(prop) %>% 
  as_flextable(show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>%
  color(color = "white", part = "header") %>%
  bold(i=6, bold=T, part="body") %>%
  italic(i=6, italic=T, part="body") %>%
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(1, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("", "Days", "Months", "Days", "Months")) %>%
  align(align = "center", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header") %>% 
  bg(i=2:6, bg = "gray80", part = "body") %>% 
  hline(i=1, border=officer::fp_border(style = "solid", width = 1.5), part="body") %>% 
  set_table_properties(layout='autofit', width=1,
                       word_description="Generalized additive models (spline models). 
                                          EDF = effective degrees of freedom. Values 
                                          close to 1 mean that linear models would fit 
                                          the data well. Values close to the number of 
                                          groups/data points mean that there might
                                          be overfitting. As EDF is a summary 
                                          statistic rather than a coefficient, 
                                          no confidence interval is presented. P-values for 
                                          the model are determined by the F-test.")

## Outputting the QQplots for the gam models
fitspline_Ad_qq = gg_qq(residuals(IgA_spline))
fitspline_Gd_qq = gg_qq(residuals(IgA_spline_m))
fitspline_Am_qq = gg_qq(residuals(IgG_spline))
fitspline_Gm_qq = gg_qq(residuals(IgG_spline_m))

plot_grid(fitspline_Ad_qq, fitspline_Gd_qq, fitspline_Am_qq, fitspline_Gm_qq, ncol=2, nrow=2, labels="AUTO") 

```

\newpage
## Other seasonal demarcations
These ways of splitting the seasons of the year were not used, but were explored in initial data analysis

```{r desc-alternative, fig.width=8, fig.height=4.5, warning=F}

# Rotavirus high and low season from VIDA
##  Low, high, shoulder season ----

data = data %>% 
  filter(IgATiter>8) %>% 
  mutate(RVseason = ifelse(month %in% c(2:4), "High (Feb-Apr)", ifelse(month %in% c(8:11), "Low (Aug-Nov)", 
                           ifelse(month %in% c(12, 1), "Intermediate (Dec, Jan)", "Intermediate (May-Jul)")))) %>% 
  mutate(RVseason = factor(RVseason, levels = c("High (Feb-Apr)",
                                                "Intermediate (May-Jul)",
                                                "Low (Aug-Nov)", 
                                                "Intermediate (Dec, Jan)")))

GMT_summaries2 = data %>% 
  group_by(RVseason) %>% 
  dplyr::summarize(N = n(), 
                   IgA_med = ci_string_dec(quantile(IgATiter, c(0.5, 0.25, 0.75)), 0), 
                   IgA_GMT = ci_string_dec(c(GMT(IgATiter), GMT_CI_se(IgATiter)), 0), 
                   IgG_med = ci_string_dec(quantile(IgGTiter, c(0.5, 0.25, 0.75)), 0), 
                   IgG_GMT = ci_string_dec(c(GMT(IgGTiter), GMT_CI_se(IgGTiter)), 0))
# shoulder seems lower than low season, suggesting there is a delay in the titer patterns, which makes sense...

GMT_summaries2 %>% 
  as_flextable(max_row = 12, show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  # colformat_double() %>% autofit() %>% 
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(2, 2, 2), top = TRUE) %>% 
  set_header_labels(values = c("Season", "N", "GMT (95% CI)", "Median (IQR)", "GMT (95% CI)", "Median (IQR)")) %>%
  align(align = "center", part = "header") %>% 
  hline(i=1:2, border=officer::fp_border(style = "solid", width = 1.5), part="header")

my_comparisons_primary = list( c("High (Feb-Apr)", "Low (Aug-Nov)"))

BP_LHS_IgA = data %>%
  filter(RVseason %in% c("High (Feb-Apr)", "Low (Aug-Nov)")) %>% 
  ggplot(aes(x=RVseason, y=IgATiter)) +
  geom_dotplot(binaxis="y",stackdir='center',binwidth = 0.05, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) +
  scale_y_log10(limits=c(10,20000), labels = scales::comma) +
  ylab("Ig (U/mL, log-scale)") + xlab(NULL) + ggtitle("IgA by Season") +
  theme_bw() + 
  ggpubr::stat_compare_means(comparisons = my_comparisons_primary, method="wilcox")

BP_LHS_IgG = data %>% 
  filter(RVseason %in% c("High (Feb-Apr)", "Low (Aug-Nov)")) %>% 
  ggplot(aes(x=RVseason, y=IgGTiter)) +
  geom_dotplot(binaxis="y",stackdir='center',binwidth = 0.05, fill="gray50", color=NA) +
  geom_boxplot(fill=NA) +
  scale_y_log10(limits=c(10,20000), labels = scales::comma) +
  ylab(NULL) + xlab(NULL) + ggtitle("IgG by Season") +
  theme_bw() + 
  ggpubr::stat_compare_means(comparisons = my_comparisons_primary, method="wilcox")

plot_grid(BP_LHS_IgA, BP_LHS_IgG, ncol=2)

ggsave("./binary_seasons.png", height = 3, width = 6, units="in", dpi=300, bg=NULL)

IgAseason_log.aov <- anova(lm(log10(IgATiter) ~ factor(RVseason), data = data))
#IgA_log.aov %>% as_flextable(show_coltype=F)

fit1Aseason = lm(log10(IgATiter) ~ factor(RVseason), data = data)

names(fit1Aseason$coefficients) = c("Intercept", "Intermediate (May-Jul)", 
                                    "Low (Aug-Nov)", "Intermediate (Dec, Jan)")

tmp_coeff = summary(fit1Aseason)$coefficients
tmp = tmp_coeff %>% as.data.frame() %>%
  mutate(est = 10^(Estimate),
         lci = 10^(Estimate-1.96*`Std. Error`), 
         uci = 10^(Estimate+1.96*`Std. Error`)) %>% 
  mutate(est_ci = paste0(my_format_mod(est, 2, 3), " (", 
                         my_format_mod(lci, 2, 3), ", ", my_format_mod(uci, 2, 3), ")")) %>% 
  mutate(pstar = ifelse(`Pr(>|t|)`<0.001, "***", 
                        ifelse(`Pr(>|t|)`<0.01, "**", 
                               ifelse(`Pr(>|t|)`<0.05, "*", ""))))

fit1Apretty = data.frame(Seasons = c("Intercept", "High (Feb-Apr)", "Intermediate (May-Jul)", 
                              "Low (Aug-Nov)", "Intermediate (Dec, Jan)"), 
          Est_A = c(tmp$est_ci[1], "Ref.", tmp$est_ci[2:4]), 
          P_A = c(tmp$pstar[1], "", tmp$pstar[2:4]))

tmp_1A_sum = summary(fit1Aseason)
tmp_1A_sum_F = tmp_1A_sum$fstatistic["value"]
tmp_1A_sum_numdf = tmp_1A_sum$fstatistic["numdf"]
tmp_1A_sum_dendf = tmp_1A_sum$fstatistic["dendf"]

tmp_ft_A = data.frame(Seasons = c("N. obs", "SE Residuals", "R2; Adj-R2", "F-stat", "P-value"), 
                      Est_A = c(length(fit1Aseason$fitted.values), 
                                my_format(sd(fit1Aseason$residuals), 2),
                                paste0(my_format(summary(fit1Aseason)$r.squared, 2), "; ", my_format(summary(fit1Aseason)$adj.r.squared, 2)),
                                my_format(tmp_1A_sum_F, 2),
                                my_smallvalue(pf(tmp_1A_sum_F, tmp_1A_sum_numdf, tmp_1A_sum_dendf, lower.tail = F), 3)))


# tmp_ft_A = fit1A %>% as_flextable() %>%
#   delete_part(part="body") %>% 
#   delete_part(part="header") 

tmp_pretty_A = fit1Apretty %>% as_flextable(max_row = 13, show_coltype=F) %>% 
  delete_rows(i = 1, part = "footer") %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header")

# vtmp_pretty_A$footer = tmp_ft_A$footer

## IgG ------
IgGseason_log.aov = anova(lm(log10(IgGTiter) ~ factor(RVseason), data = data))

# lm(log2(IgGTiter) ~ factor(month, levels=c(10, 1:9, 11:12)), data = data) %>% summary
fit1Gseason = lm(log10(IgGTiter) ~ factor(RVseason), data = data) # %>% summary

names(fit1Gseason$coefficients) = c("Intercept", "Intermediate (May-Jul)", 
                              "Low (Aug-Nov)", "Intermediate (Dec, Jan)")

tmp_coeff = summary(fit1Gseason)$coefficients
tmp = tmp_coeff %>% as.data.frame() %>%
  mutate(est = 10^(Estimate),
         lci = 10^(Estimate-1.96*`Std. Error`), 
         uci = 10^(Estimate+1.96*`Std. Error`)) %>% 
  mutate(est_ci = paste0(my_format_mod(est, 2, 3), " (", 
                my_format_mod(lci, 2, 3), ", ", my_format_mod(uci, 2, 3), ")")) %>% 
  mutate(pstar = ifelse(`Pr(>|t|)`<0.001, "***", 
                ifelse(`Pr(>|t|)`<0.01, "**", 
                ifelse(`Pr(>|t|)`<0.05, "*", ""))))

fit1Gpretty = data.frame(Seasons = c("Intercept", "High (Feb-Apr)", "Intermediate (May-Jul)", 
                              "Low (Aug-Nov)", "Intermediate (Dec, Jan)"), 
          Est_G = c(tmp$est_ci[1], "Ref.", tmp$est_ci[2:4]), 
          P_G = c(tmp$pstar[1], "", tmp$pstar[2:4]))

tmp_1G_sum = summary(fit1Gseason)
tmp_1G_sum_F = tmp_1G_sum$fstatistic["value"]
# tmp_1G_sum_numdf = tmp_1G_sum$fstatistic["numdf"]
# tmp_1G_sum_dendf = tmp_1G_sum$fstatistic["dendf"]

tmp_ft_G = data.frame(Seasons = c("N. obs", "SE Residuals", "R2; Adj-R2", "F-stat", "P-value"), 
                      Est_G = c(length(fit1Gseason$fitted.values), 
                                my_format(sd(fit1Gseason$residuals), 2),
                                paste0(my_smallvalue(summary(fit1Gseason)$r.squared, 2), "; ", my_smallvalue(summary(fit1Gseason)$adj.r.squared, 2)),
                                my_format(tmp_1G_sum_F, 2),
                                my_smallvalue(fpvalue(fit1Gseason)$Fp, 3)))
# pf(tmp_1G_sum_F, tmp_1G_sum_numdf, tmp_1G_sum_dendf, lower.tail = F)

fit1Apretty %>% 
  bind_rows(tibble(tmp_ft_A)) %>%
  full_join(fit1Gpretty %>% bind_rows(tmp_ft_G)) %>%
  as_flextable(max_row = 30, show_coltype=F, spread_first_col = TRUE) %>% 
  delete_part(part = "footer") %>%
  bg(bg = "black", part = "header") %>% 
  color(color = "white", part = "header") %>% 
  bg(i=6:10, bg = "gray80", part = "body") %>% 
  hline(i=5, border=officer::fp_border(style = "solid", width = 2), part="body") %>%
  add_header_row(values = c("", "IgA", "IgG"), colwidths = c(1, 2, 2), top = TRUE) %>% 
  hline(i=1, border=officer::fp_border(style = "solid", width = 2), part="header") %>% 
  # align(align = "center", part = "header") %>% 
  line_spacing(space = 1, part = "all") %>% 
  add_footer_lines("Signif. codes: 0 ≤ '***' < 0.001 < '**' < 0.01 < '*' < 0.05") %>% 
  italic(part="footer", italic=T) # %>% 

```
